Drought vulnerability of tree growth

### 2024-07-08 scripts/explore_analysis.Rmd
updated script

@@

png("growth-precip-thailand/doc/display/census_increment_species.png", width=4, height=4, units="in", res=300)
ggplot(hkk_census2 %>% filter(sp %in% c("SACCLI", "HOPEOD", "TETRNU", "POLYVI")), 
aes(x= factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y = avg_inc, fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))) +
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
  scale_fill_viridis_d()+
  labs(title = 'tree growth in HKK from census', 
  y="Annualised diameter increment (cm)", x="species") +
  theme_bw()+
  #geom_vline(xintercept=0, lty=2)+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()
```

@@

spagplot_all<-ggplot() +
  #species plots
  geom_line(data=dendro_inc %>% group_by(Cno, sp) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc, 
  group=sp, col=sp))+
  #mean of all trees
  geom_line(data=dendro_inc %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="black", size=2)+
  #mean of species
  geom_line(data=dendro_inc %>% group_by(Cno, sp) %>% 
  dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)) %>%
  ungroup()%>%
  group_by(Cno) %>% dplyr::summarise(mean_inc=mean(mean_inc, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="grey40", size=0.8)+
  scale_color_viridis_d()+
  guides(col="none", group="none")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 4 species")+
  theme_bw()

@@

#sample size per species per year
png("growth-precip-thailand/doc/display/top4_sample_size.png", width=8, height=4, units="in", res=300)
top4 %>% 
  group_by(sp, Cno) %>% 
  dplyr::summarise(n=n())%>%
  ggplot(aes(x=Cno, y=n, fill=factor(sp,levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_bar(stat="identity")+
  scale_fill_viridis_d()+
  facet_grid(.~factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))+
  xlab("census number")+
  ylab("sample size")+
  guides(fill=guide_legend(title="species"))+
  theme_bw()
dev.off()

@@

growth_top4<-ggplot(top4, aes(x=factor(((Cno-3)/2)+2009), y=inc_annual,
                                              fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
                                              color=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))
                                              ))+ 
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
ylab("Annualised diameter increment (cm)")+theme_bw()+
facet_grid(.~factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))+
  geom_hline(yintercept = 0, lty=3)+
      theme(axis.text.x = element_text(angle=90))+
  ggtitle("HKK dendroband data from 2008 to 2023")+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  guides(fill=guide_legend(title="species"),
         color=guide_legend(title="species"))+
         xlab("year")

@@

growth_top4_sum<-ggplot(top4, aes(x=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y=inc_annual,
                                              fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
                                              color=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))
                                              ))+ 
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
ylab("Annualised diameter increment (cm)")+
xlab("species")+theme_bw()+
  geom_hline(yintercept = 0, lty=3)+
  ggtitle("tree growth in HKK from dendrobands")+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  guides(fill="none",
         color="none")

@@

@@

png("growth-precip-thailand/doc/display/resistance_map_top4.png", width=10, height=8, units="in", res=300)
resistance_map
dev.off()

>>>>

census_top4<-bind_rows(hkk.stem1, hkk.stem2, hkk.stem3, hkk.stem4) %>%
  filter(status=="A")%>%
  filter(sp %in% c("SACCLI", "HOPEOD", "TETRNU", "POLYVI"))%>%
  filter(tag==StemTag) %>%
  group_by(sp, tag)%>%
  dplyr::mutate(dbhmin1 = dplyr::lag(dbh, n=1, default=NA, order_by=tag),
                inc_annual = (dbh-dbhmin1)/5)%>%
  filter(!is.na(dbhmin1))%>%
  #average increment in cm
  dplyr::summarise(avg_inc = mean(inc_annual/10, na.rm=T),
                   sd_inc = sd(inc_annual/10, na.rm=T))

@@

png("growth-precip-thailand/doc/display/census_increment_species.png", width=4, height=4, units="in", res=300)
ggplot(hkk_census2 %>% filter(sp %in% c("SACCLI", "HOPEOD", "TETRNU", "POLYVI")), 
aes(x= factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y = avg_inc, fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))) +
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
  scale_fill_viridis_d()+
  labs(title = 'tree growth in HKK from census', 
  y="Annualised diameter increment (cm)", x="species") +
  theme_bw()+
  geom_text(data=hkk_census2 %>% filter(sp %in% c("SACCLI", "HOPEOD", "TETRNU", "POLYVI")) %>% group_by(sp) %>% dplyr::summarise(n=n()),
   aes(x=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y=2.2, 
  label=n))+
  #geom_vline(xintercept=0, lty=2)+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()

png("growth-precip-thailand/doc/display/census_increment_species_alltrees.png", width=4, height=4, units="in", res=300)
ggplot(census_top4, 
aes(x= factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y = avg_inc, fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))) +
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
  scale_fill_viridis_d()+
  labs(title = 'tree growth in HKK (all stems)', 
  y="Annualised diameter increment (cm)", x="species") +
  theme_bw()+
  geom_text(data=census_top4 %>% group_by(sp) %>% dplyr::summarise(n=n()),
   aes(x=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y=2.2, 
  label=n))+
  #geom_vline(xintercept=0, lty=2)+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
dev.off()

@@

#spaghetti plot for top 10 species

sp_top10<-c("SACCLI", "HOPEOD", "TETRNU", "POLYVI", "VATIHA", 
"DIPTAL", "ALPHVE", "GARCSP", "GARUPI", "NEOLOB")

spagplot_top10<-ggplot()+
  #species plots
  geom_line(data=dendro_inc %>% filter(sp %in% sp_top10)%>% group_by(Cno, sp) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc, 
  group=sp, col=sp))+
  #mean of all trees
  geom_line(data=dendro_inc %>% filter(sp %in% sp_top10)%>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="black", size=2)+
  #mean of species
  geom_line(data=dendro_inc %>% filter(sp %in% sp_top10)%>% group_by(Cno, sp) %>% 
  dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)) %>%
  ungroup()%>%
  group_by(Cno) %>% dplyr::summarise(mean_inc=mean(mean_inc, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="grey40", size=0.8)+
  scale_color_viridis_d()+
  guides(col=guide_legend("species"))+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 10 species")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_top10.png", width=4, height=4, units="in", res=300)
spagplot_top10
dev.off()

spagplot_all<-ggplot() +
  #species plots
  geom_line(data=dendro_inc %>% group_by(Cno, sp) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc, 
  group=sp, col=sp))+
  #mean of all trees
  geom_line(data=dendro_inc %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="black", size=2)+
  #mean of species
  geom_line(data=dendro_inc %>% group_by(Cno, sp) %>% 
  dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)) %>%
  ungroup()%>%
  group_by(Cno) %>% dplyr::summarise(mean_inc=mean(mean_inc, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="grey40", size=0.8)+
  scale_color_viridis_d()+
  guides(col="none", group="none")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for all species")+
  theme_bw()

@@

@@

growth_top4<-ggplot(top4, aes(x=factor(((Cno-3)/2)+2009), y=inc_annual,
                                              fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
                                              color=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))
                                              ))+ 
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
ylab("Annualised diameter increment (cm)")+theme_bw()+
geom_segment(data=top4 %>% 
  group_by(sp, Cno) %>% 
  dplyr::summarise(n=n()/500),
aes(x=factor(((Cno-3)/2)+2009), xend=factor(((Cno-3)/2)+2009), y=-3, yend=n-3, 
fill=factor(sp,levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))), size=3)+
facet_grid(.~factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")))+
  geom_hline(yintercept = 0, lty=3)+
      theme(axis.text.x = element_text(angle=90))+
  ggtitle("HKK dendroband data from 2008 to 2023")+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  guides(fill=guide_legend(title="species"),
         color=guide_legend(title="species"))+
         xlab("year")

@@

growth_top4_sum<-ggplot(top4, aes(x=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y=inc_annual,
                                              fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
                                              color=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))
                                              ))+ 
  geom_violin(position = "dodge")+ 
    geom_boxplot(position="dodge", col="black", fill="white", width=0.1)+
ylab("Annualised diameter increment (cm)")+
xlab("species")+theme_bw()+
  geom_text(data=top4 %>% group_by(sp) %>% dplyr::summarise(n=n()),
   aes(x=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), y=2.2, 
  label=n))+
  geom_hline(yintercept = 0, lty=3)+
  ggtitle("tree growth in HKK from dendrobands")+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  guides(fill="none",
         color="none")

@@

```
```{r top4 species growth contribution, echo=F}

#calculate contribution of each species to average growth for each year

top4_sum<-top4 %>%
  group_by(Cno) %>%
  dplyr::mutate(tot_inc_all=sum(inc_annual, na.rm=T))%>%
  ungroup()%>%
  group_by(Cno, sp) %>%
  dplyr::summarise(contrib=sum(inc_annual, na.rm=T)/mean(tot_inc_all))

#plot contributions across years

contrib_top4<-ggplot(top4_sum, 
aes(x=factor(((Cno-3)/2)+2009), 
y=contrib, fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_bar(stat="identity")+
  scale_fill_viridis_d()+
  xlab("year")+
  ylab("contribution to total growth")+
  guides(fill=guide_legend(title="species"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90))

png("growth-precip-thailand/doc/display/top4_contributions.png", width=4, height=4, units="in", res=300)
contrib_top4
dev.off()

#repeat contributions plot for all species

species_sum <- dendro_inc %>%
  group_by(Cno) %>%
  dplyr::mutate(tot_inc_all=sum(inc_annual, na.rm=T))%>%
  ungroup()%>%
  group_by(Cno, sp) %>%
  dplyr::summarise(contrib=sum(inc_annual, na.rm=T)/mean(tot_inc_all))

#plot contributions across years

contrib_all<-ggplot(species_sum, 
aes(x=factor(((Cno-3)/2)+2009), 
y=contrib, fill=factor(sp)))+
  geom_bar(stat="identity")+
  scale_fill_viridis_d()+
  xlab("year")+
  ylab("contribution to total growth")+
  guides(fill=guide_legend(title="species"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90))

png("growth-precip-thailand/doc/display/species_contributions.png", width=12, height=4, units="in", res=300)
contrib_all
dev.off()

@@

png("growth-precip-thailand/doc/display/resistance_map_top4.png", width=9, height=8, units="in", res=300)
resistance_map
dev.off()
